---
title: "Assignment 1"
author: "Natalie Alexander, Thabo Dube & Kelly-Robyn Singh"
date: "2023-10-06"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(stringr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(textdata)
library(tidytext)
library(gridExtra)

```

# 1. Reading in the data

```{r Reading in Data}

set.seed(2022)

sona = readRDS("C:/Users/natal/OneDrive - University of Cape Town/UCT/Data_science_for_industry/Group_Assignment/STA5073Z_Assignment2/Natalie/preprocessed_sona.rds") # you can change to where you have it saved

sona<-sona%>% mutate(speech= str_replace(speech, "\\d{1,2} [A-Za-z]+ \\d{4}", "")) # Remove dates at the start of the speech

sona<- sona%>% mutate(speech= str_replace(speech, pattern = "^Thursday, ", replacement = ""))# remove dates on 2 remaining Ramaphosa speeches

#remove white spaces
sona<- sona%>% mutate(speech= str_trim(speech, side= "left"))

#tokenize to sentences
Sona_S_tokenized<- unnest_tokens(sona, sentence, speech, token = 'sentences') 

#remove punctuation 
Sona_S_tokenized <- Sona_S_tokenized%>% mutate(sentence= str_replace_all(sentence, "[[:punct:]]", ""))

```

# 2. Sentiment Lexicon

```{r Data analyisis} 

afinn <- get_sentiments('afinn') # Lexicon of sentiments with value attached ranging from -5 to 5 

```

# 3. Bigram tokenization

```{r}

#tokenize sentences to words
df_bigrams <- Sona_S_tokenized %>%
  unnest_tokens(bigram, sentence, token = 'ngrams', n = 2, to_lower = T) %>%
select(bigram, everything())
  
#separate the bigrams  into individual words
df_separated_bigrams <- df_bigrams %>%
  separate(bigram, c('word1', 'word2'), sep = ' ')

# remove stop words
df_filtered_bigrams <- df_separated_bigrams %>%
  filter(!word1 %in% stop_words$word & !word2 %in% stop_words$word) 


```

#  4. Sentiment analysis on bigrams using the Afinn lexicon

```{r}

#use afinn sentiments 
##join with sentiment data based on word1
bigrams_word1_sentiment <- df_separated_bigrams %>%
  select(word1)%>%
  left_join(afinn, by = c("word1" = "word")) %>%
  rename(word1_sentiment = value)

##join with sentiment data based on word2
bigrams_word2_sentiment <- df_separated_bigrams %>%
  select(word2)%>%
  left_join(afinn, by = c("word2" = "word")) %>%
  rename(word2_sentiment = value)


#remove word 1 and word 2 column from data frame
df_separated_bigrams = df_separated_bigrams %>% select(filename, year, president_13, date)

#get NETT sentiment for each bigram and append as a column to the data frame
bigrams_sentiment = cbind(df_separated_bigrams, bigrams_word1_sentiment , bigrams_word2_sentiment )
 
################################################################################

#convert NA to 0
bigrams_sentiment$word1_sentiment[is.na(bigrams_sentiment$word1_sentiment)] = 0 #word1
bigrams_sentiment$word2_sentiment[is.na(bigrams_sentiment$word2_sentiment)] = 0 #word2

#extract sentiments for word1 and word2 and calculate total bigram sentiment
total_sentiment = rowSums(cbind(bigrams_sentiment$word1_sentiment, bigrams_sentiment$word2_sentiment))

#append total sentiment to bigrams data frame
df_bigram_total_sentiment = bigrams_sentiment %>% select(word1, word2, filename, year, president_13, date)%>%
  mutate(sentiment=total_sentiment)

###############################################################################

#negative, if nett sentiment < 0
#positive, else if nett sentiment > 0
#neutral if nett sentiment = 0

total_logical_sentiment  = ifelse(df_bigram_total_sentiment$sentiment < 0, "negative", ifelse(df_bigram_total_sentiment$sentiment > 0, "positive", "neutral"))
                                                                                                  
#replace total numeric sentiment with categorical sentiment
df_bigram_total_sentiment = df_bigram_total_sentiment %>%
  mutate(sentiment=total_logical_sentiment)

###############################################################################

#Find negation words

##negation words
negation_words <- c('not', 'no', 'never', 'without')

#show negated bigrams
filter(df_bigram_total_sentiment, word1 %in% negation_words) %>% 
  head(10) %>% select(word1, word2, sentiment, president_13, year)
  
#convert sentiment to positive if word 1 in negation_words and sentiment = negative
#convert sentiment to negative if word 1 in negation_words and sentiment = positive
#else, neutral.
df_bigram_total_sentiment = df_bigram_total_sentiment %>%
  mutate(sentiment = ifelse(word1 %in% negation_words & sentiment == "positive", "negative",
                            ifelse(word1 %in% negation_words & sentiment == "negative", "positive", sentiment)))

# join up the bigrams again
df_bigrams <- df_bigram_total_sentiment %>%
  unite(bigram, word1, word2, sep = ' ')

#remove neutral sentiment
df_bigrams  = df_bigrams %>% select(filename, bigram, sentiment, president_13, year) %>% filter(sentiment == "negative" | sentiment == "positive")

```


## 4.1 Results of sentiment analysis per president

```{r}


#sentiment analysis per president

## empty vector to append NETT sentiment
nett_president_sentiment = c()

## deKlerk numeric sentiment
deklerk_sentiment = df_bigrams %>% filter(president_13=="deKlerk")

#find majority sentiment
sentiment_mat = as.matrix(table(deklerk_sentiment$sentiment))
rownames(sentiment_mat)[which.max(sentiment_mat)]

#append sentiment to list
nett_president_sentiment = c(nett_president_sentiment, sentiment)

###############################################################################

## Mandela numeric sentiment
mandela_sentiment = df_bigrams %>% filter(president_13=="Mandela")

#find majority sentiment
sentiment_mat = as.matrix(table(mandela_sentiment$sentiment))
rownames(sentiment_mat)[which.max(sentiment_mat)]

#append sentiment to list
nett_president_sentiment = c(nett_president_sentiment, sentiment)

###############################################################################

## Mbeki numeric sentiment
mbeki_sentiment = df_bigrams %>% filter(president_13=="Mbeki")

#find majority sentiment
sentiment_mat = as.matrix(table(mbeki_sentiment$sentiment))
rownames(sentiment_mat)[which.max(sentiment_mat)]

#append sentiment to list
nett_president_sentiment = c(nett_president_sentiment, sentiment)

###############################################################################

## Motlanthe numeric sentiment
motlanthe_sentiment = df_bigrams %>% filter(president_13=="Motlanthe")

#find majority sentiment
sentiment_mat = as.matrix(table(motlanthe_sentiment$sentiment))
rownames(sentiment_mat)[which.max(sentiment_mat)]

#append sentiment to list
nett_president_sentiment = c(nett_president_sentiment, sentiment)

###############################################################################


## Zuma numeric sentiment
zuma_sentiment = df_bigrams %>% filter(president_13=="Zuma")

#find majority sentiment
sentiment_mat = as.matrix(table(zuma_sentiment$sentiment))
rownames(sentiment_mat)[which.max(sentiment_mat)]

#append sentiment to list
nett_president_sentiment = c(nett_president_sentiment, sentiment)

###############################################################################


## Ramaphosa numeric sentiment
ramaphosa_sentiment = df_bigrams %>% filter(president_13=="Ramaphosa")

#find majority sentiment
sentiment_mat = as.matrix(table(ramaphosa_sentiment$sentiment))
sentiment = rownames(sentiment_mat)[which.max(sentiment_mat)]

#append sentiment to list
nett_president_sentiment = c(nett_president_sentiment, sentiment)

#set president names to vector
names(nett_president_sentiment) = c("deklerk", "mandela", "mbeki", "motlanthe",
                                    "zuma", "ramaphosa")

```

## 4.2 Results of sentiment analysis per speech

```{r}


#file names of each unique file containing a speech
filenames = unique(df_bigram_total_sentiment$filename)

#find speeches with negative sentiment
for (speech in filenames) {
  
  speech_sentiment = df_bigrams %>% filter(filename == speech)
  
  #table of sentiment counts
  sentiment_mat = as.matrix(table(speech_sentiment$sentiment))
  
  # Check if sentiment_mat has any values
  if (length(sentiment_mat) > 0) {
    sentiment = rownames(sentiment_mat)[which.max(sentiment_mat)] #max sentiment

    # Check if sentiment is not null before using it in an if statement
    if (!is.null(sentiment) && sentiment == "negative") {  
      print(speech) #print speeches with negative sentiment
    }
  }
}

```

## 4.3 Results of sentiment analysis per year

```{r}

#years
years = unique(df_bigram_total_sentiment$year)

#find yeares with negative sentiment
for (yr in years) {
  
  year_sentiment = df_bigrams %>% filter(year == yr)
  
  #table of sentiment counts
  sentiment_mat = as.matrix(table(year_sentiment$sentiment))
  
  # Check if sentiment_mat has any values
  if (length(sentiment_mat) > 0) {
    sentiment = rownames(sentiment_mat)[which.max(sentiment_mat)] #max sentiment
    
    # Check if sentiment is not null before using it in an if statement
    if (!is.null(sentiment) && sentiment == "negative") {  
      print(yr) #print year with negative sentiment
    }
  }
}

```

## 4.4. Plot top 5 bigrams related to negative and positive sentiment for each president

```{r}

####################negative

##Mandela: Top 5 bigrams related to negative sentiments
p1 = df_bigrams %>% filter(sentiment == "negative" & president_13 == "Mandela")%>%
  count(bigram, sort = TRUE) %>% 
    filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("c. Mandela: Top 5 negative bigrams")+
  theme(legend.position = "none")



##Mbeki: Top 5 bigrams related to negative sentiments
p2 = df_bigrams %>% filter(sentiment == "negative" & president_13 == "Mbeki")%>%
  count(bigram, sort = TRUE) %>% 
    filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("e. Mbeki: Top 5 negative bigrams")+
  theme(legend.position = "none")



##Ramaphosa: Top 5 bigrams related to negative sentiments
p3 = df_bigrams %>% filter(sentiment == "negative" & president_13 == "Ramaphosa")%>%
  count(bigram, sort = TRUE) %>%
  filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("k. Ramaphosa: Top 5 negative bigrams")+
  theme(legend.position = "none")



##Zuma: Top 5 bigrams related to negative sentiments
p4 = df_bigrams %>% filter(sentiment == "negative" & president_13 == "Zuma")%>%
  count(bigram, sort = TRUE) %>%   filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("i. Zuma: Top 5 negative bigrams")+
  theme(legend.position = "none")



##deKlerk: Top 5 bigrams related to negative sentiments
p5 = df_bigrams %>% filter(sentiment == "negative" & president_13 == "deKlerk")%>%
  count(bigram, sort = TRUE)  %>% 
    filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("a. deKlerk: Top 5 negative bigrams")+
  theme(legend.position = "none")



##Motlanthe: Top 5 bigrams related to negative sentiments
p6 = df_bigrams %>% filter(sentiment == "negative" & president_13 == "Motlanthe")%>%
  count(bigram, sort = TRUE) %>% 
    filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("g. Motlanthe: Top 5 negative bigrams")+
  theme(legend.position = "none")


####################positive

##Mandela: Top 5 bigrams related to positive sentiments
p7 = df_bigrams %>% filter(sentiment == "positive" & president_13 == "Mandela")%>%
  count(bigram, sort = TRUE) %>% 
    filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("d. Mandela: Top 5 positive bigrams")+
  theme(legend.position = "none")



##Mbeki: Top 5 bigrams related to positive sentiments
p8 = df_bigrams %>% filter(sentiment == "positive" & president_13 == "Mbeki")%>%
  count(bigram, sort = TRUE) %>% 
    filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("f. Mbeki: Top 5 positive bigrams")+
  theme(legend.position = "none")



##Ramaphosa: Top 5 bigrams related to positive sentiments
p9 = df_bigrams %>% filter(sentiment == "positive" & president_13 == "Ramaphosa")%>%
  count(bigram, sort = TRUE) %>% 
    filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("l. Ramaphosa: Top 5 positive bigrams")+
  theme(legend.position = "none")



##Zuma: Top 5 bigrams related to positive sentiments
p10 = df_bigrams %>% filter(sentiment == "positive" & president_13 == "Zuma")%>%
  count(bigram, sort = TRUE) %>% 
    filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("j. Zuma: Top 5 positive bigrams")+
  theme(legend.position = "none")



##deKlerk: Top 5 bigrams related to positive sentiments
p11 = df_bigrams %>% filter(sentiment == "positive" & president_13 == "deKlerk")%>%
  count(bigram, sort = TRUE) %>% 
    filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("b. deKlerk: Top 5 positive bigrams")+
  theme(legend.position = "none")



##Motlanthe: Top 5 bigrams related to positive sentiments
p12 = df_bigrams %>% filter(sentiment == "positive" & president_13 == "Motlanthe")%>%
  count(bigram, sort = TRUE)%>%
  filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("h. Motlanthe: Top 5 positive bigrams")+
  theme(legend.position = "none")

###############grid plot
grid.arrange(p5, p11,
              p1, p7,  
             p2, p8, 
              p6, p12,
              p4, p10,
               p3, p9,
               ncol = 2)

```
Figure #: Bar plots showing the top 5 bigrams with the highest occurring frequency per president and sentiment class, where the top 5 ranked bigrams may not necessarily mean 5 bars.

## 4.5 Plot top 5 bigrams related to negative and positive sentiments across all years and for all speeches and presidents

```{r}

##Top 5 bigrams related to negative sentiments
p1 = df_bigrams %>% filter(sentiment == "negative")%>%
  count(bigram, sort = TRUE) %>%
  filter(rank(desc(n)) <= 5) %>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("a. Top 5 negative bigrams across all presidents")+
  theme(legend.position = "none")


##Top 5 bigrams related to positive sentiments
p2 = df_bigrams %>% filter(sentiment == "positive")%>%
  count(bigram, sort = TRUE) %>%
  filter(rank(desc(n)) <= 5)%>%
  arrange(desc(n)) %>%
  ggplot(aes(reorder(bigram,n),n, fill=bigram)) + 
  geom_col() + coord_flip() + xlab('')+
  ggtitle("b. Top 5 positive bigrams across all presidents")+
  theme(legend.position = "none")

###############grid plot
grid.arrange(p1, p2, ncol=2)

```
Figure #: Barplots showing results of LDA topic modelling using beta (bigram-topic) probabilities and k = 2 topics.

# 5. LDA Topic modelling

## 5.1 LDA Topic modelling for all presidents

```{r}

library(topicmodels)

#frequency of each bigram in each document
sona_tdf <- df_bigrams %>%
  group_by(filename,  bigram) %>%
  count() %>%  
  ungroup() 

#Reshape long format to wide format 
#put data frame into the DocumentTermMatrix class 
dtm_sona <- sona_tdf %>% 
  cast_dtm(filename, bigram, n)

#run LDA
sona_lda <- LDA(dtm_sona, k = 2, control = list(seed = 123)) #choose different values of k to find best value of k
str(sona_lda)

#beta-bigram topic probabilities
sona_topics <- tidy(sona_lda, matrix = 'beta')
#head(sona_topics)

#Plot top 20 terms used in each topic
sona_topics %>%
  group_by(topic) %>%
  filter(!term %in% c("south africans", "south african", "south africas", #remove south africa root words
                      "south africa", "fellow south", "madame speaker",
                      "speaker chairperson", "madam speaker")) %>% #remove unnecessary words
  slice_max(n = 20, order_by = beta) %>% ungroup() %>%
  arrange(topic, -beta) %>%
  ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = 'free') + coord_flip()

###############

```
Figure #: Barplots showing results of LDA topic modelling using beta (bigram-topic) probabilities and k = 2 topics.

## 5.2.  LDA Topic modelling for each president

### 5.2.1 de Klerk

```{r}
#LDA Topic Modelling for deKlerk

#frequency of each bigram in each document
sona_tdf_deKlerk <- df_bigrams %>%
  filter(president_13=="deKlerk")%>%
  group_by(filename,  bigram) %>%
  count() %>%  
  ungroup() 

#Reshape long format to wide format 
#put data frame into the DocumentTermMatrix class 
dtm_sona_deKlerk <- sona_tdf_deKlerk%>% 
  cast_dtm(filename, bigram, n)

#run LDA
sona_lda_deKlerk <- LDA(dtm_sona_deKlerk, k = 2, control = list(seed = 123)) #choose different values of k to find best value of k
str(sona_lda_deKlerk)

#beta-bigram topic probabilities
sona_topics_deKlerk <- tidy(sona_lda_deKlerk, matrix = 'beta')
#head(sona_topics_deKlerk)

#Plot top 20 terms used in each topic
# sona_topics_deKlerk %>%
#   group_by(topic) %>%
#   filter(!term %in% c("south africans", "south african", "south africas", #remove south africa root words
#                       "south africa", "fellow south", "madame speaker",
#                       "speaker chairperson", "madam speaker")) %>% #remove unnecessary words
#   slice_max(n = 20, order_by = beta) %>% ungroup() %>%
#   arrange(topic, -beta) %>%
#   ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~ topic, scales = 'free') + coord_flip()

##############

##  which words have the greatest difference in beta values between Topic 1 and Topic 2.
beta_spread <- sona_topics_deKlerk %>%
  mutate(topic = paste0('topic', topic)) %>%
  pivot_wider(names_from = topic, values_from = beta) %>%
  filter(topic1 > .001 | topic2 > .001) %>%
  mutate(log_ratio = log2(topic2 / topic1))

p1 = beta_spread %>%
  group_by(direction = log_ratio > 0) %>%
  top_n(10, abs(log_ratio)) %>%
  ungroup() %>%
  mutate(term = reorder(term, log_ratio)) %>%
  ggplot(aes(term, log_ratio, fill=direction)) +
  geom_col() +
  labs(y = '') +
  coord_flip()+
  ggtitle("a. deKlerk: beta values for topic 1 vs topic 2")+
  theme(legend.position = "none")

```

### 5.2.2 Mandela

```{r}
#LDA Topic Modelling for each president:

###########

#LDA Topic Modelling for Mandela

#frequency of each bigram in each document
sona_tdf_mandela <- df_bigrams %>%
  filter(president_13=="Mandela")%>%
  group_by(filename,  bigram) %>%
  count() %>%  
  ungroup() 

#Reshape long format to wide format 
#put data frame into the DocumentTermMatrix class 
dtm_sona_mandela <- sona_tdf_mandela %>% 
  cast_dtm(filename, bigram, n)

#run LDA
sona_lda_mandela <- LDA(dtm_sona_mandela, k = 2, control = list(seed = 123)) #choose different values of k to find best value of k
str(sona_lda_mandela)

#beta-bigram topic probabilities
sona_topics_mandela <- tidy(sona_lda_mandela, matrix = 'beta')
#head(sona_topics_mandela)

#Plot top 20 terms used in each topic
  # sona_topics_mandela %>%
  #   group_by(topic) %>%
  #   filter(!term %in% c("south africans", "south african", "south africas", #remove south africa root words
  #                       "south africa", "fellow south", "madame speaker",
  #                       "speaker chairperson", "madam speaker")) %>% #remove unnecessary words
  #   slice_max(n = 20, order_by = beta) %>% ungroup() %>%
  #   arrange(topic, -beta) %>%
  #   ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
  #   geom_col(show.legend = FALSE) +
  #   facet_wrap(~ topic, scales = 'free') + coord_flip()+
  #   ggtitle("Mandela: LDA Topic Model")

##############

##  which words have the greatest difference in beta values between Topic 1 and Topic 2.
beta_spread <- sona_topics_mandela %>%
  mutate(topic = paste0('topic', topic)) %>%
  pivot_wider(names_from = topic, values_from = beta) %>%
  filter(topic1 > .001 | topic2 > .001) %>%
  mutate(log_ratio = log2(topic2 / topic1))

p2 = beta_spread %>%
  group_by(direction = log_ratio > 0) %>%
  top_n(10, abs(log_ratio)) %>%
  ungroup() %>%
  mutate(term = reorder(term, log_ratio)) %>%
  ggplot(aes(term, log_ratio, fill=direction)) +
  geom_col() +
  labs(y = '') +
  coord_flip()+
  ggtitle("b. Mandela: beta values for topic 1 vs topic 2")+
  theme(legend.position = "none")

```

### 5.2.3 Mbeki

```{r}
#LDA Topic Modelling for Mbeki

#frequency of each bigram in each document
sona_tdf_mbeki<- df_bigrams %>%
  filter(president_13=="Mbeki")%>%
  group_by(filename,  bigram) %>%
  count() %>%  
  ungroup() 

#Reshape long format to wide format 
#put data frame into the DocumentTermMatrix class 
dtm_sona_mbeki<- sona_tdf_mbeki%>% 
  cast_dtm(filename, bigram, n)

#run LDA
sona_lda_mbeki <- LDA(dtm_sona_mbeki, k = 2, control = list(seed = 123)) #choose different values of k to find best value of k
str(sona_lda_mbeki)

#beta-bigram topic probabilities
sona_topics_mbeki <- tidy(sona_lda_mbeki, matrix = 'beta')
#head(sona_topics_mbeki)

#Plot top 20 terms used in each topic
# sona_topics_mbeki %>%
#   group_by(topic) %>%
#   filter(!term %in% c("south africans", "south african", "south africas", #remove south africa root words
#                       "south africa", "fellow south", "madame speaker",
#                       "speaker chairperson", "madam speaker")) %>% #remove unnecessary words
#   slice_max(n = 20, order_by = beta) %>% ungroup() %>%
#   arrange(topic, -beta) %>%
#   ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~ topic, scales = 'free') + coord_flip()+
#   ggtitle("Mbeki: LDA Topic Model")

##############

##  which words have the greatest difference in beta values between Topic 1 and Topic 2.
beta_spread <- sona_topics_mbeki %>%
  mutate(topic = paste0('topic', topic)) %>%
  pivot_wider(names_from = topic, values_from = beta) %>%
  filter(topic1 > .001 | topic2 > .001) %>%
  mutate(log_ratio = log2(topic2 / topic1))

p3 = beta_spread %>%
  group_by(direction = log_ratio > 0) %>%
  top_n(10, abs(log_ratio)) %>%
  ungroup() %>%
  mutate(term = reorder(term, log_ratio)) %>%
  ggplot(aes(term, log_ratio, fill=direction)) +
  geom_col() +
  labs(y = '') +
  coord_flip()+
  ggtitle("c. Mbeki: beta values for topic 1 vs topic 2")+
  theme(legend.position = "none")

```

### 5.2.4 Motlanthe

```{r}
#LDA Topic Modelling for Motlanthe

#frequency of each bigram in each document
sona_tdf_motlanthe <- df_bigrams %>%
  filter(president_13=="Motlanthe")%>%
  group_by(filename,  bigram) %>%
  count() %>%  
  ungroup() 

#Reshape long format to wide format 
#put data frame into the DocumentTermMatrix class 
dtm_sona_motlanthe <- sona_tdf_motlanthe%>% 
  cast_dtm(filename, bigram, n)

#run LDA
sona_lda_motlanthe <- LDA(dtm_sona_motlanthe, k = 2, control = list(seed = 123)) #choose different values of k to find best value of k
str(sona_lda_motlanthe)

#beta-bigram topic probabilities
sona_topics_motlanthe <- tidy(sona_lda_motlanthe, matrix = 'beta')
#head(sona_topics_motlanthe)

#Plot top 20 terms used in each topic
# sona_topics_motlanthe %>%
#   group_by(topic) %>%
#   filter(!term %in% c("south africans", "south african", "south africas", #remove south africa root words
#                       "south africa", "fellow south", "madame speaker",
#                       "speaker chairperson", "madam speaker")) %>% #remove unnecessary words
#   slice_max(n = 20, order_by = beta) %>% ungroup() %>%
#   arrange(topic, -beta) %>%
#   ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~ topic, scales = 'free') + coord_flip()


##############

##  which words have the greatest difference in beta values between Topic 1 and Topic 2.
beta_spread <- sona_topics_motlanthe %>%
  mutate(topic = paste0('topic', topic)) %>%
  pivot_wider(names_from = topic, values_from = beta) %>%
  filter(topic1 > .001 | topic2 > .001) %>%
  mutate(log_ratio = log2(topic2 / topic1))

p4 = beta_spread %>%
  group_by(direction = log_ratio > 0) %>%
  top_n(10, abs(log_ratio)) %>%
  ungroup() %>%
  mutate(term = reorder(term, log_ratio)) %>%
  ggplot(aes(term, log_ratio, fill=direction)) +
  geom_col() +
  labs(y = '') +
  coord_flip()+
  ggtitle("d. Motlanthe: beta values for topic 1 vs topic 2")+
  theme(legend.position = "none")

##########
```


### 5.2.5 Zuma

```{r}
#LDA Topic Modelling for Zuma

#frequency of each bigram in each document
sona_tdf_zuma <- df_bigrams %>%
  filter(president_13=="Zuma")%>%
  group_by(filename,  bigram) %>%
  count() %>%  
  ungroup() 

#Reshape long format to wide format 
#put data frame into the DocumentTermMatrix class 
dtm_sona_zuma <- sona_tdf_zuma%>% 
  cast_dtm(filename, bigram, n)

#run LDA
sona_lda_zuma <- LDA(dtm_sona_zuma, k = 2, control = list(seed = 123)) #choose different values of k to find best value of k
str(sona_lda_zuma)

#beta-bigram topic probabilities
sona_topics_zuma <- tidy(sona_lda_zuma, matrix = 'beta')
#head(sona_topics_zuma)

#Plot top 20 terms used in each topic
# sona_topics_zuma %>%
#   group_by(topic) %>%
#   filter(!term %in% c("south africans", "south african", "south africas", #remove south africa root words
#                       "south africa", "fellow south", "madame speaker",
#                       "speaker chairperson", "madam speaker")) %>% #remove unnecessary words
#   slice_max(n = 20, order_by = beta) %>% ungroup() %>%
#   arrange(topic, -beta) %>%
#   ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~ topic, scales = 'free') + coord_flip()+
#   ggtitle("Zuma: LDA Topic Model")

##############

##  which words have the greatest difference in beta values between Topic 1 and Topic 2.
beta_spread <- sona_topics_zuma %>%
  mutate(topic = paste0('topic', topic)) %>%
  pivot_wider(names_from = topic, values_from = beta) %>%
  filter(topic1 > .001 | topic2 > .001) %>%
  mutate(log_ratio = log2(topic2 / topic1))

p5 = beta_spread %>%
  group_by(direction = log_ratio > 0) %>%
  top_n(10, abs(log_ratio)) %>%
  ungroup() %>%
  mutate(term = reorder(term, log_ratio)) %>%
  ggplot(aes(term, log_ratio, fill=direction)) +
  geom_col() +
  labs(y = 'Log2 ratio of beta in topic 2 / topic 1') +
  coord_flip()+
  ggtitle("e. Zuma: beta values for topic 1 vs topic 2")+
  theme(legend.position = "none")

###########
```



### 5.2.6 Ramaphosa

```{r}
###########

#LDA Topic Modelling for Ramaphosa

#frequency of each bigram in each document
sona_tdf_rama <- df_bigrams %>%
  filter(president_13=="Ramaphosa")%>%
  group_by(filename,  bigram) %>%
  count() %>%  
  ungroup() 

#Reshape long format to wide format 
#put data frame into the DocumentTermMatrix class 
dtm_sona_rama<- sona_tdf_rama%>% 
  cast_dtm(filename, bigram, n)

#run LDA
sona_lda_rama <- LDA(dtm_sona_rama, k = 2, control = list(seed = 123)) #choose different values of k to find best value of k
str(sona_lda_rama)

#beta-bigram topic probabilities
sona_topics_rama <- tidy(sona_lda_rama, matrix = 'beta')
#head(sona_topics_rama)

#Plot top 20 terms used in each topic
# sona_topics_rama %>%
#   group_by(topic) %>%
#   filter(!term %in% c("south africans", "south african", "south africas", #remove south africa root words
#                       "south africa", "fellow south", "madame speaker",
#                       "speaker chairperson", "madam speaker")) %>% #remove unnecessary words
#   slice_max(n = 20, order_by = beta) %>% ungroup() %>%
#   arrange(topic, -beta) %>%
#   ggplot(aes(reorder(term, beta), beta, fill = factor(topic))) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~ topic, scales = 'free') + coord_flip()+
#   ggtitle("Ramaphosa: LDA Topic Model")

##############

##  which words have the greatest difference in beta values between Topic 1 and Topic 2.
beta_spread <- sona_topics_rama %>%
  mutate(topic = paste0('topic', topic)) %>%
  pivot_wider(names_from = topic, values_from = beta) %>%
  filter(topic1 > .001 | topic2 > .001) %>%
  mutate(log_ratio = log2(topic2 / topic1))

p6 = beta_spread %>%
  group_by(direction = log_ratio > 0) %>%
  top_n(10, abs(log_ratio)) %>%
  ungroup() %>%
  mutate(term = reorder(term, log_ratio)) %>%
  ggplot(aes(term, log_ratio, fill=direction)) +
  geom_col() +
  labs(y = '') +
  coord_flip()+
  ggtitle("f. Ramaphosa: beta values for topic 1 vs topic 2")+
  theme(legend.position = "none")

###########
```

### 5.2.7. Grid plot

```{r}

###############grid plot
grid.arrange(p1, p2,
             p3, p4, 
             p5, p6,
               ncol = 3)
```
Figure #: Barplots showing results of LDA topic modelling using beta (bigram-topic) probabilities and k = 2 topics for each former president class. Topic 1 is shown in red and topic 2 is shown in green.

